Linearizability Proofs Based On Forward Simulations

### 2017-01-19 draftCAV/queue.tex

Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation. The linearization point of a dequeue operation with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before order stored in the current state contains a minimal enqueue operation that adds the value $d$. The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state. When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores no enqueue operation. The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point.

>>>>

Informally, $AbsQ$ records the happens-before order between enqueue operations for which the added value has not been removed by a dequeue operation. The linearization point of a dequeue operation with return value $d\neq{\tt EMPTY}$ is enabled only if the happens-before order stored in the current state contains a minimal enqueue operation that adds the value $d$. The effect of the linearization point is that the minimal enqueue is removed from the current state and the return value is recorded in the library state. When the return value is {\tt EMPTY}, the linearization point of a dequeue is enabled only if the current state stores no enqueue operation. The return of a dequeue is enabled only if the returned value matches the one fixed at the linearization point. \textcolor{red}{Important: I think EMPTY return is problematic and I defined it wrong in my machine too. We should be able to return EMPTY if there are only pending nodes. Consider the following history of $AbsQ_0$: $inv(enq, d_1,k_1), inv(enq, d_2,k_2), inv(deq, k_3), lin(deq, \texttt{EMPTY}, k_3)$. This history should be reflected in $AbsQ$ by enabling lin \texttt{EMPTY} of dequeue when there are pending nodes. We also need to update the rules in figure.}

### 2017-01-19 draftCAV/stack.tex

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$. We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.}
  \label{fig:transitions:AbsS}
\vspace{-6mm}
\end{figure}

>>>>

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$. We use the following notions: $f[\forall k.guard(k)=> k\mapsto expr(k)]$ is the function $g$ defined by $g(k)=f(k)$ for all $k$ such that $guard(k)$ is false, and $g(k)=expr(k)$ otherwise, $maxCo(O)$ is the set of maximally completed operations in $O$, i.e., $maxCo(O)=\set{k\in O: \ell_1(k)={\tt COMP}, \forall k'\in O.\ k' < k \vee \ell_1(k')={\tt PEND}}$, ${\tt PEND}(O)=\{k\in O: \ell_2(k)={\tt PEND}\}$, and $pred_{<}(k')$ is the set of immediate predecessors of $k'$ according to $<$, i.e., $pred_{<}(k')=\set{k\in O: k < k'\land \forall k''\in O.\ k'' > k' \vee k'' < k}$.\textcolor{red}{call-push should have $d \neq EMPTY$ in the premise. If I understand correctly, there is a problem change in $be$ in lin-pop1 rule. What I understand is we move $k''$ to every predecesoor of $k'$. But we should not move it to predecessor of $k'$ if this predecessor has a successor $m$ such that $m \in be(k'')$. Is this case reflected in your definition. Please read sub-bullet 4 of bullet 4 in the definition of $L_I$ in my document. Lin-pop2 seems ok for me. Last note: Bibliography seems not correct to me.} }
  \label{fig:transitions:AbsS}
\vspace{-6mm}
\end{figure}
